---
title: "harmonization"
output: html_document
date: "2026-02-06"
---
# ============================================================================
# NbS STOCKTAKE - DATA HARMONIZATION TO TYPOLOGY
# ============================================================================
# 
# PURPOSE: Map extracted free-text values to standardized typology categories
# 
# WORKFLOW:
#   PART 1: Setup & Load Results
#   PART 2: Define Harmonization Prompts
#   PART 3: Run Harmonization
#   PART 4: Save Harmonized Results
#
# ============================================================================

# Load libraries
```{r libraries}
library(dplyr)
library(glue)
library(readr)
library(writexl)
library(ellmer)
library(DT)
```

# ============================================================================
# PART 1: CONFIGURATION & LOAD DATA
# ============================================================================
```{r config}
# ----- API Configuration -----
openai_api_key <- Sys.getenv("NBS_API_KEY")
GPT_MODEL <- "gpt-5-nano"

# ----- Paths -----
# Input: Results from extraction
INPUT_RESULTS <- "C:/PDFs_extracted_text/extraction_results/nbs_test_extraction_results2.rds"

# Output directory
OUT_DIR <- "C:/PDFs_extracted_text/harmonization_results"
dir.create(OUT_DIR, showWarnings = FALSE, recursive = TRUE)

# ----- Helper Functions -----
`%||%` <- function(a, b) if (!is.null(a) && !is.na(a) && nzchar(a)) a else b

# Safe API call wrapper
safe_chat <- function(prompt, model = GPT_MODEL) {
  tryCatch({
    chat <- ellmer::chat_openai(model = model)
    resp <- chat$chat(prompt)
    list(response = trimws(resp), error = NA_character_)
  }, error = function(e) {
    list(response = NA_character_, error = conditionMessage(e))
  })
}
```

```{r load_data}
# Load extraction results
if (file.exists(INPUT_RESULTS)) {
  extraction_results <- readRDS(INPUT_RESULTS)
  message(glue("Loaded {nrow(extraction_results)} papers for harmonization"))
} else {
  stop(glue("Input file not found: {INPUT_RESULTS}"))
}
```

# ============================================================================
# PART 2: HARMONIZATION PROMPTS
# ============================================================================
```{r harmonization_prompts}
# ----- Prompt: Harmonize Input Variables -----
harmonize_input_variables <- function(raw_value) {
  glue('
You are harmonizing extracted data to a standardized typology.

RAW EXTRACTED VALUE:
"""
{raw_value}
"""

Map each variable/data layer mentioned above to ONE of these standardized categories from the GEE Awesome Catalog typology:

1. Population and Socioeconomic Datasets
2. Hydrology Datasets
3. Global Land Use and Land Cover Datasets
4. Climate and Weather Datasets
5. Elevation and terrain
6. Soil properties and geology
7. Biodiversity, ecosystems and habitat
8. Earth observations imagery and derived indices
9. Build environment and infrastructure
10. Hazard disturbance and events
11. Governance management and institutional indicators

RULES:
- Map each variable to the MOST appropriate category
- If a variable fits multiple categories, choose the PRIMARY category
- Format: Category name; Category name; Category name
- Keep the order of variables as they appear in the raw value
- Use exact category names from the list above
- Output ONLY the category names separated by semicolons
- If the raw value is NA or no variables are present, return NA

EXAMPLES:
Input: "Slope; Rainfall; Population density; NDVI"
Output: Elevation and terrain; Climate and Weather Datasets; Population and Socioeconomic Datasets; Earth observations imagery and derived indices

Input: "DEM; Soil pH; Distance to roads"
Output: Elevation and terrain; Soil properties and geology; Build environment and infrastructure

Output only the harmonized categories on a single line. No explanation.')
}

# ----- Prompt: Harmonize Input Variable Resolution -----
harmonize_input_resolution <- function(raw_value) {
  glue('
You are harmonizing extracted data to a standardized typology.

RAW EXTRACTED VALUE:
"""
{raw_value}
"""

Map the resolution(s) to ONE of these standardized categories:

1. Not reported
2. Non gridded / admin / qualitative
3. High: <30m
4. Medium: 30-250m
5. Coarse: 250m-1km
6. Very coarse: >1km


RULES:
- If multiple resolutions with different categories are present, list them separated by semicolumns
- If resolution mentioned is administrative/vector/qualitative, use "Non gridded / admin / qualitative"
- Examples:
  * "30m" → High: <30m
  * "100m" → Medium: 30-250m
  * "1km" → Coarse: 250m-1km
  * "10km" → Very coarse: >1km
  * "county-level" → Non gridded / admin / qualitative
  * "not specified" → Not reported
- Use exact category names from the list above
- Output ONLY ONE category name
- If raw value is NA, return "Not reported"

Output only the harmonized category on a single line. No explanation.')
}

# ----- Prompt: Harmonize Spatial Method -----
harmonize_spatial_method <- function(raw_value) {
  glue('
You are harmonizing extracted data to a standardized typology.

RAW EXTRACTED VALUE:
"""
{raw_value}
"""

Map the spatial method(s) to ONE or MORE of these standardized categories:

1. MCDA / MCE: Weighted overlay, WLC, AHP, OWA, fuzzy MCDA
2. Machine learning prediction: RF, SVM, XGBoost, ANN
3. Process based biophysical models: hydrology/erosion/ecosystem services models
4. Spatial optimization: Marxan/Zonation, simulated annealing, ILP
5. Spatial statistics / hotspots analysis
6. Land change scenarios models
7. Rule based GIS Overlay and constraints: Buffers, distance thresholds, masks
8. Bayesian probabilistic DSS
9. Participatory/ expert driven DSS

RULES:
- Map each method to the MOST appropriate category
- A paper can use multiple method categories - list all that apply separated by semicolumns
- Format: Category name; Category name
- Examples:
  * "AHP" → MCDA / MCE
  * "Random Forest" → Machine learning prediction
  * "Buffer analysis" → Rule based GIS Overlay
  * "InVEST" → Process based biophysical models
  * "AHP and Random Forest" → MCDA / MCE; Machine learning prediction
- Use exact category names from the list above
- If raw value is NA or no method identified, return NA

Output only the harmonized category/categories on a single line. No explanation.')
}

# ----- Prompt: Harmonize Method Validation -----
harmonize_method_validation <- function(raw_value) {
  glue('
You are harmonizing extracted data to a standardized typology.

RAW EXTRACTED VALUE:
"""
{raw_value}
"""

Map the validation approach(es) to ONE or MORE of these standardized categories:

1. Field verification/ground-truthing
2. Comparison with existing NbS sites or known locations
3. Comparison with independent datasets
4. Cross-validation (k-fold, leave-one-out)
5. Split-sample validation (training/test data)
6. Expert review or stakeholder validation
7. Comparison with other models or studies
8. Sensitivity analysis
9. Uncertainty analysis

RULES:
- Map each validation approach to the MOST appropriate category
- Multiple validation approaches can be present - list all that apply separated by semicolumns
- Format: Category name; Category name
- Use exact category names from the list above
- If raw value is NA or no validation mentioned, return NA

Output only the harmonized category/categories on a single line. No explanation.')
}

# ----- Prompt: Harmonize Analysis Resolution -----
harmonize_analysis_resolution <- function(raw_value) {
  glue('
You are harmonizing extracted data to a standardized typology.

RAW EXTRACTED VALUE:
"""
{raw_value}
"""

Map the analysis resolution to ONE of these standardized categories:

1. Not reported
2. Raster cell size (high: <30m)
3. Raster cell size (medium: 30-250m)
4. Raster cell size (coarse: 250m-1km)
5. Raster cell size (very coarse: >1km)
6. Regular grid
7. Irregular planning units
8. Administrative units
9. Hydrological units

RULES:
- Choose the SINGLE most appropriate category
- If multiple resolutions at different scales, choose the PRIMARY analysis resolution
- Examples:
  * "30m resolution" → Raster cell size (high: <30m)
  * "100m pixels" → Raster cell size (medium: 30-250m)
  * "1km grid" → Raster cell size (coarse: 250m-1km)
  * "10km x 10km grid" → Raster cell size (very coarse: >1km)
  * "county-level" → Administrative units
  * "watershed-based" → Hydrological units
  * "hexagon grid 5km spacing" → Regular grid
- Use exact category names from the list above
- If raw value is NA or not specified, return "Not reported"

Output only ONE harmonized category on a single line. No explanation.')
}

# ----- Prompt: Harmonize Output -----
harmonize_output <- function(raw_value) {
  glue('
You are harmonizing extracted data to a standardized typology.

RAW EXTRACTED VALUE:
"""
{raw_value}
"""

Map the output(s) to ONE or MORE of these standardized categories:

1. Suitability / feasibility maps
2. Priority / ranking maps
3. Vulnerability / risk / hazard maps
4. Ecosystem service / biophysical quantification
5. Conservation network / selection solutions: Spatial optimization outputs
6. Scenarios comparisons / trade-offs
7. Economic outputs
8. Non-Spatial decision outputs: MCS strategy ranking; indicator tables

RULES:
- Map each output to the MOST appropriate category
- Multiple output types can be present - list all that apply separated by semicolumnns
- Format: Category name; Category name
- Examples:
  * "Suitability map for agroforestry" → Suitability / feasibility maps
  * "Priority areas ranked by score" → Priority / ranking maps
  * "Flood risk zones" → Vulnerability / risk / hazard maps
  * "Carbon sequestration estimates" → Ecosystem service / biophysical quantification
  * "Cost-benefit ratios" → Economic outputs
- Use exact category names from the list above
- If raw value is NA, return NA

Output only the harmonized category/categories on a single line. No explanation.')
}

# ----- Prompt: Harmonize Climate Risk Assessment -----
harmonize_climate_risk <- function(raw_value) {
  glue('
You are harmonizing extracted data to a standardized typology.

RAW EXTRACTED VALUE:
"""
{raw_value}
"""

Map the climate risk approach to ONE of these standardized categories:

1. None
2. Baseline climate predictors: Uses temperature/precip/bioclim variables but not framed as hazard/vulnerability
3. Historical hazard/event layers: Uses observed/historical floods/fires/storms etc as hazards
4. Variability/trend/probabilistic climate: Uses time series trends drought indices, exceedance probabilities
5. Future climate/SLR scenarios: Explicit projections
6. Full risk framework: Explicit hazard + exposure + vulnerability

RULES:
- Choose the SINGLE most comprehensive/advanced category that applies
- Hierarchy (if multiple apply, choose highest):
  * Full risk framework > Future scenarios > Variability/trends > Historical hazards > Baseline climate > None
- Examples:
  * "No" → None
  * "Uses temperature and rainfall as inputs" → Baseline climate predictors
  * "Historical flood zones used" → Historical hazard/event layers
  * "RCP 4.5 and 8.5 scenarios" → Future climate/SLR scenarios
  * "Vulnerability assessment with hazard, exposure, and adaptive capacity" → Full risk framework
- Use exact category names from the list above

Output only ONE harmonized category on a single line. No explanation.')
}

# ----- Prompt: Harmonize Economics -----
harmonize_economics <- function(raw_value) {
  glue('
You are harmonizing extracted data to a standardized typology.

RAW EXTRACTED VALUE:
"""
{raw_value}
"""

Map the economic analysis to ONE of these standardized categories:

1. None
2. Cost mapping: Spatially explicit cost estimates are produced but no aggregation comparison or valuation of benefits
3. Economic feasibility / cost comparison: Costs are compared, ranked or used to assess feasibility
4. Monetized ecosystem service valuation
5. Cost benefit analysis
6. Net present value
7. Cost effectiveness optimization

RULES:
- Choose the SINGLE most comprehensive/advanced category that applies
- Hierarchy (if multiple apply, choose highest):
  * Cost effectiveness optimization > Net present value > Cost benefit analysis > Monetized ES valuation > Economic feasibility > Cost mapping > None
- Examples:
  * "No" → None
  * "Implementation costs mapped" → Cost mapping
  * "Costs compared across regions" → Economic feasibility / cost comparison
  * "Ecosystem services valued in dollars" → Monetized ecosystem service valuation
  * "NPV calculated for 20-year period" → Net present value
  * "Benefit-cost ratio computed" → Cost benefit analysis
- Use exact category names from the list above

Output only ONE harmonized category on a single line. No explanation.')
}

# ----- Prompt: Harmonize Geographic Scope -----
harmonize_geographic_scope <- function(raw_value) {
  glue('
You are harmonizing extracted data to a standardized typology.

RAW EXTRACTED VALUE:
"""
{raw_value}
"""

Map the geographic scope to ONE of these continental categories:

1. Africa
2. Europe
3. North America
4. South America
5. Oceania
6. Asia
7. Global / Multicontinental

RULES:
- Extract ONLY the continental/regional category
- If multiple countries from same continent, use that continent
- If multiple countries from different continents, use "Global / Multicontinental"
- Examples:
  * "Sub-national: Ethiopia" → Africa
  * "National: Kenya" → Africa
  * "Multi-country: South Asia" → Asia
  * "Sub-national: Iceland" → Europe
  * "National: USA" → North America
  * "Multi-country: Kenya and India" → Global / Multicontinental
- Use exact category names from the list above
- If raw value is NA, return NA

Output only ONE harmonized category on a single line. No explanation.')
}
```

# ============================================================================
# PART 3: HARMONIZATION FUNCTIONS
# ============================================================================
```{r harmonization_functions}
# ----- Single variable harmonizer -----
harmonize_single_variable <- function(results_df, raw_column, 
                                     harmonize_fn, harmonized_column_name) {
  
  n_papers <- nrow(results_df)
  harmonized <- vector("character", n_papers)
  
  pb <- txtProgressBar(min = 0, max = n_papers, style = 3)
  
  for (i in seq_len(n_papers)) {
    raw_value <- results_df[[raw_column]][i] %||% "NA"
    
    # Skip if already NA
    if (is.na(raw_value) || raw_value == "NA") {
      harmonized[i] <- NA_character_
      setTxtProgressBar(pb, i)
      next
    }
    
    # Build prompt
    prompt <- harmonize_fn(raw_value)
    
    # Call API
    resp <- safe_chat(prompt)
    
    if (!is.na(resp$error)) {
      harmonized[i] <- NA_character_
      warning(glue("Error harmonizing {harmonized_column_name} for paper {i}: {resp$error}"))
    } else {
      harmonized[i] <- resp$response
    }
    
    setTxtProgressBar(pb, i)
    Sys.sleep(0.5)  # Rate limiting
  }
  
  close(pb)
  harmonized
}
```

# ============================================================================
# PART 4: RUN HARMONIZATION
# ============================================================================
```{r run_harmonization, eval=FALSE}
# Initialize harmonized results with original data
harmonized_results <- extraction_results

# ----- Harmonize each variable -----

# === Input Variables ===
message("\n=== Harmonizing Input Variables ===")
harmonized_results$Input_Variables_Harmonized <- harmonize_single_variable(
  harmonized_results,
  raw_column = "Input_Variables",
  harmonize_fn = harmonize_input_variables,
  harmonized_column_name = "Input_Variables_Harmonized"
)

# === Input Variable Resolution ===
message("\n=== Harmonizing Input Variable Resolution ===")
harmonized_results$Input_Variable_Resolution_Harmonized <- harmonize_single_variable(
  harmonized_results,
  raw_column = "Input_Variable_Resolution",
  harmonize_fn = harmonize_input_resolution,
  harmonized_column_name = "Input_Variable_Resolution_Harmonized"
)

# === Spatial Method ===
message("\n=== Harmonizing Spatial Method ===")
harmonized_results$Spatial_Method_Harmonized <- harmonize_single_variable(
  harmonized_results,
  raw_column = "Spatial_Method",
  harmonize_fn = harmonize_spatial_method,
  harmonized_column_name = "Spatial_Method_Harmonized"
)

# === Method Validation ===
message("\n=== Harmonizing Method Validation ===")
harmonized_results$Method_Validation_Harmonized <- harmonize_single_variable(
  harmonized_results,
  raw_column = "Method_Validation",
  harmonize_fn = harmonize_method_validation,
  harmonized_column_name = "Method_Validation_Harmonized"
)

# === Analysis Resolution ===
message("\n=== Harmonizing Analysis Resolution ===")
harmonized_results$Analysis_Resolution_Harmonized <- harmonize_single_variable(
  harmonized_results,
  raw_column = "Analysis_Resolution",
  harmonize_fn = harmonize_analysis_resolution,
  harmonized_column_name = "Analysis_Resolution_Harmonized"
)

# === Output ===
message("\n=== Harmonizing Output ===")
harmonized_results$Output_Harmonized <- harmonize_single_variable(
  harmonized_results,
  raw_column = "Output",
  harmonize_fn = harmonize_output,
  harmonized_column_name = "Output_Harmonized"
)

# === Climate Risk Assessment ===
message("\n=== Harmonizing Climate Risk Assessment ===")
harmonized_results$Climate_Risk_Assessment_Harmonized <- harmonize_single_variable(
  harmonized_results,
  raw_column = "Climate_Risk_Assessment",
  harmonize_fn = harmonize_climate_risk,
  harmonized_column_name = "Climate_Risk_Assessment_Harmonized"
)

# === Economics ===
message("\n=== Harmonizing Economics ===")
harmonized_results$Economics_Harmonized <- harmonize_single_variable(
  harmonized_results,
  raw_column = "Economics",
  harmonize_fn = harmonize_economics,
  harmonized_column_name = "Economics_Harmonized"
)

# === Geographic Scope ===
message("\n=== Harmonizing Geographic Scope ===")
harmonized_results$Geographic_Scope_Harmonized <- harmonize_single_variable(
  harmonized_results,
  raw_column = "Geographic_Scope",
  harmonize_fn = harmonize_geographic_scope,
  harmonized_column_name = "Geographic_Scope_Harmonized"
)

message("\n✓ Harmonization complete!")
```

# ============================================================================
# PART 5: SAVE HARMONIZED RESULTS
# ============================================================================
```{r save_harmonized, eval=FALSE}
# Save as RDS
saveRDS(harmonized_results, file.path(OUT_DIR, "nbs_harmonized_results.rds"))

# Save as CSV
write_csv(harmonized_results, file.path(OUT_DIR, "nbs_harmonized_results.csv"))


message(glue("\nHarmonized results saved to: {OUT_DIR}"))
message(glue("  - nbs_harmonized_results.rds"))
message(glue("  - nbs_harmonized_results.csv"))

# ----- Summary statistics -----
harmonization_summary <- harmonized_results %>%
  summarise(
    Total_Papers = n(),
    Input_Vars_Harmonized = sum(!is.na(Input_Variables_Harmonized)),
    Spatial_Method_Harmonized = sum(!is.na(Spatial_Method_Harmonized)),
    Output_Harmonized = sum(!is.na(Output_Harmonized)),
    Climate_Risk_Harmonized = sum(!is.na(Climate_Risk_Assessment_Harmonized)),
    Economics_Harmonized = sum(!is.na(Economics_Harmonized)),
    Geographic_Scope_Harmonized = sum(!is.na(Geographic_Scope_Harmonized))
  )

print(harmonization_summary)
```

# ============================================================================
# PART 6: COMPARISON VIEW (Original vs Harmonized)
# ============================================================================
```{r comparison_view, eval=FALSE}
# Create side-by-side comparison for review
comparison_df <- harmonized_results %>%
  select(
    File,
    NbS_Practice,
    # Input Variables
    Input_Variables,
    Input_Variables_Harmonized,
    # Spatial Method
    Spatial_Method,
    Spatial_Method_Harmonized,
    # Output
    Output,
    Output_Harmonized,
    # Climate Risk
    Climate_Risk_Assessment,
    Climate_Risk_Assessment_Harmonized,
    # Economics
    Economics,
    Economics_Harmonized,
    # Geographic Scope
    Geographic_Scope,
    Geographic_Scope_Harmonized
  )

# Interactive comparison table
DT::datatable(
  comparison_df %>%
    mutate(across(where(is.character), ~ifelse(
      nchar(.) > 60,
      paste0(substr(., 1, 60), "..."),
      .
    ))),
  options = list(pageLength = 10, scrollX = TRUE),
  caption = "Original vs Harmonized Values - Comparison View"
)
```

```{r}
write_csv(harmonized_results, file.path(OUT_DIR, "test_harmonization_results.csv"))
message(glue("\nTest results saved to: {OUT_DIR}/test_harmonization_results.csv"))
```


